---
title: "23 April 2019 - Anime"
author: "Alyssa Vanderbeek"
date: "4/23/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)

theme_set(theme_bw())
```

```{r}
tidy_anime <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-23/tidy_anime.csv") %>%
  dplyr::select(-synopsis, -background)
```

## Exploratory analysis

```{r}
# number of genres for each show 
tidy_anime %>%
  dplyr::select(animeID, genre) %>%
  group_by(animeID) %>%
  count

# number of shows for each category 
tidy_anime %>%
  dplyr::select(animeID, genre) %>%
  group_by(genre) %>%
  count %>%
  arrange(desc(n))

# average score for each genre
genre_score = tidy_anime %>%
  dplyr::select(genre, score) %>%
  group_by(genre) %>%
  summarise(mean = mean(score, na.rm = T)) %>%
  arrange(desc(mean)) %>%
  dplyr::select(genre) %>%
  unlist
```


Some questions I might in interested in exploring are:
What is the distribution of scores for a given genre?
Do shows that were on the air for longer have higher scores than newer shows?
Do the most popular shows have the highest ratings? Importantly, what's the relationship between scores and number of users who scored?


## Distribution of scores in a genre
```{r}
tidy_anime %>%
  dplyr::select(genre, score, popularity) %>%
  filter(genre != "NA") %>%
  group_by(genre) %>%
  ggplot(aes(y = score, fill = genre)) +
  geom_boxplot() 
```


## Relationship between length of time on air and score
```{r}
tidy_anime %>%
  mutate(length_time = as.duration(start_date %--% end_date) / dyears(1)) %>%
  filter(genre %in% genre_score[1:5]) %>%
  dplyr::select(animeID, score, length_time, genre, members) %>%
  ggplot(aes(x = length_time, y = score, size = members, color = genre)) +
  geom_point(alpha = 0.3) +
  facet_grid(genre ~ .)
  scale_x_continuous(limits = c(0, 10))
```


## Relationship between show popularity, show score, and number of users who scored
```{r}

tidy_anime %>%
  mutate(year = as.numeric(format(start_date, '%Y'))) %>% # release year
  filter(year > 1960) %>% # Most shows with dates earlier than this have missing data
  mutate(year = cut(year,
                    breaks = c(1960, 1970, 1980, 1990, 2000, 2010, Inf),
                    labels = c("1960s", "1970s", "1980s", "1990s", "2000s", "2010s"))) %>% # group year into decades
  dplyr::select(animeID, year, score, scored_by, members) %>% # select relevant cols
  unique() %>% # one row for one show
  ggplot(aes(x = members, y = score, size = scored_by)) +
  geom_point(alpha = 0.3) +
  facet_grid(year ~ .) +
  coord_flip() + 
  labs(
    title = 'Anime popularity, score, and number of scorers',
    y = 'Score',
    x = 'Number of members with the anime in their list',
    size = 'Number of users who\nhave scored the show'  
    ) +
  theme(legend.position = "bottom") 

```

Initially in the above plot I was using `popularity` as a measure of popularity (it seemed like a reasonable thing to do). But I was finding that there was a negative correlation with the number of users who scored the anime, such that low values of "popularity" were scored by a large number of people. This might have made sense if the scores were low, but in fact this phenomenon was more prevalent in the higher scoring shows. Instead, I switched my measure of popularity to the `members` variable, which gives a much more intuitive measure of popularity. 

From this figure we can see that higher scores are actually associated with greater popularity and a greater number of scorers. Newer shows (after 2000) are also more popular and more highly scored than older shows.